\texttt{nuance}: Detection of planetary transits in the presence of correlated noise

### 2023-02-11 latex/ms.tex
fix again proof of period. likelihood

i.e., given a depth $D$, the likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$ and a common depth $\Delta$. Since only $\{p(\bm{f} \vert T_k, D, \Delta_k)\}_{k}$ is known (i.e. transits with different depths), we decompose
\begin{equation}\label{eq:non_part_of_per}
    p(\bm{f} \vert T_k, D, \Delta) = \int p(\bm{f} \vert T_k, D, \tilde\Delta)p(\tilde\Delta | \Delta)\, d\tilde\Delta
\end{equation}
where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$. In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
\\\\
Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood function:
$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k\exp \left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)$$
\\\\
As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2)$$
with 
\begin{equation}\label{eq:params}
\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \hspace{0.5cm} \text{and} \hspace{0.5cm}
\Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}
\end{equation}
Hence
$$p(\tilde\Delta | \Delta) = \frac{1}{\sqrt{2\pi\sigma^2}}x\exp \left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)$$
We can now rewrite \autoref{eq:non_part_of_per} as
\begin{equation}\label{eq:ice_int}
    p(\bm{f} \vert T_k, D, \Delta) =  \frac{\mathcal{L}_k}{\sqrt{2\pi\sigma^2}} \int \exp\left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)\, \exp\left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)\, d\tilde\Delta
\end{equation}
The integral in \autoref{eq:ice_int} is a product of gaussian integrals that can be obtained analytically and we have
\begin{proof}{gaussians_product_integral}
    p(\bm{f} \vert T_k, D, \Delta) = \mathcal{L}_k  \sqrt{\frac{2\pi\sigma_{k}^2}{\sigma^{2} + \sigma_{k}^{2}}} \exp\left(-\frac{1}{2}\frac{(\Delta_k-\Delta)^2}{\sigma_k^2 + \sigma^2}\right)
\end{proof}
And \autoref{eq:goal} gives 
\begin{equation}\label{eq:result}
    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(2\pi\sigma_{k}^2) + \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right)
\end{equation}
the log-likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and common depth $\Delta$. In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \hspace{0.5cm}\text{and}\hspace{0.5cm} \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.

>>>>

i.e., given a depth $D$, the likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$ and a common depth $\Delta$. Since only $\{p(\bm{f} \vert T_k, D, \Delta_k)\}_{k}$ is known (i.e. transits with different depths), we decompose
\begin{equation}\label{eq:non_part_of_per}
    p(\bm{f} \vert T_k, D, \Delta) = \int p(\bm{f} \vert T_k, D, \tilde\Delta)p(\tilde\Delta | \Delta)\, d\tilde\Delta
\end{equation}
where $p(\bm{f} \vert T_k, D, \tilde\Delta)$ is the probability of the $k$-th transit to have a depth $\tilde\Delta$ and $p(\tilde\Delta | \Delta)$ the probability to observe the depth $\tilde\Delta$ knowing the existence of a common depth $\Delta$. In other words, \autoref{eq:non_part_of_per} involves the likelihood of the non-periodic transit $k$ to be part of a periodic transit signal with a common depth $\Delta$.
\\\\
Since each depth $\Delta_k$ is found through generalized least square, each follow a normal distribution $\mathcal{N}(\Delta_k, \sigma_k^2)$, centered on $\Delta_k$ with variance $\sigma_k^2$, and with an amplitude $\mathcal{L}_k$, leading to the likelihood function:
$$p(\bm{f} \vert T_k, D, \tilde\Delta) = \mathcal{L}_k\exp \left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)$$
\\\\
As for the common transit depth $\Delta$, it can be estimated through the joint probability of all other transit depths than $\Delta_k$, such that
$$\Delta \sim \prod_{\sumTk}^K \mathcal{N}(\Delta_i, \sigma_i^2)$$
with 
\begin{equation}\label{eq:params}
\frac{1}{\sigma^2} = \sum_{\sumTk}^K \frac{1}{\sigma_i^2} \hspace{0.5cm} \text{and} \hspace{0.5cm}
\Delta =\sigma^2 \sum_{\sumTk}^K {\frac{\Delta_i}{\sigma_i^2}}
\end{equation}
Hence
$$p(\tilde\Delta | \Delta) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp \left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)$$
We can now rewrite \autoref{eq:non_part_of_per} as
\begin{equation}\label{eq:ice_int}
    p(\bm{f} \vert T_k, D, \Delta) =  \frac{\mathcal{L}_k}{\sqrt{2\pi\sigma^2}} \int \exp\left(-\frac{(\tilde\Delta-\Delta_k)^2}{2\sigma_k^2}\right)\, \exp\left(-\frac{(\tilde\Delta-\Delta)^2}{2\sigma^2}\right)\, d\tilde\Delta
\end{equation}
The integral in \autoref{eq:ice_int} is a product of gaussian integrals that can be obtained analytically and we have
\begin{proof}{gaussians_product_integral}
    p(\bm{f} \vert T_k, D, \Delta) = \mathcal{L}_k  \sqrt{\frac{\sigma_{k}^2}{\sigma^{2} + \sigma_{k}^{2}}} \exp\left(-\frac{1}{2}\frac{(\Delta_k-\Delta)^2}{\sigma_k^2 + \sigma^2}\right)
\end{proof}
And \autoref{eq:goal} gives 
\begin{equation}\label{eq:result}
    \ln p(\bm{f} \vert P, T_0 ,D, \Delta) =  \sum_{k}^K \ln \mathcal{L}_k  - \frac{1}{2} \sum_k^K\left(\ln(\sigma_{k}^2) - \ln(\sigma^{2} + \sigma_{k}^{2}) +  \frac{\left(\Delta_{k} -
    \Delta\right)^{2}}{\sigma_k^{2} + \sigma^{2}}\right)
\end{equation}
the log-likelihood of our data given a periodic transit signal of period $P$, epoch $T_0$, duration $D$ and common depth $\Delta$. In order to reduce the number of times \autoref{eq:params} is computed, we adopt the biased estimates
$$\frac{1}{\sigma^2} = \sum_{k}^K \frac{1}{\sigma_i^2} \hspace{0.5cm}\text{and}\hspace{0.5cm} \Delta  = \sigma^2 \sum_{k}^K {\frac{\Delta_i}{\sigma_i^2}}$$
so that $\Delta$ and $\sigma$ are independent of $k$ in the last sum of \autoref{eq:result}.
